Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model \thanks{Grants or other notes about the article that should go on the front page should be placed here. General acknowledgments should be placed at the end of the article.}

### 2021-07-20 cover.tex
cover letter updated

%% comment macros
\newcommand{\irena}[1]{\textcolor{irena}{$\langle$\textbf{IP:} #1$\rangle$}}
\newcommand{\de}[1]{\textcolor{de}{$\langle$\textbf{DE:} #1$\rangle$}}

%% journal name macro
\newcommand{\journalname}{\emph{BMC Public Health}}

%% data macros
% \input{data_macros.tex}

@@

Dr. Natalie Pafitis

@@

\begin{flushright}
Irena Papst

Center for Applied Mathematics, Cornell University

657 Rhodes Hall

136 Hoy Rd, Ithaca, NY, USA 14853

\href{mailto:ip98@cornell.edu}{ip98@cornell.edu}
\end{flushright}

@@

Dear Dr.\ Pafitis,

We are pleased to submit an original research article entitled
``Age-dependence of healthcare interventions for COVID-19 in Ontario,
Canada'' for your consideration for publication in \journalname{}. This study specifically aligns with the aims of \journalname{} to publish research related to the social determinants of health and adds new insight into the evolving COVID-19 pandemic.

Age is the most salient clinical indicator of risk from COVID-19 for
most patients. While age-specific distributions of known SARS-CoV-2
infections and COVID-19-related deaths have been widely reported,
relatively few sources present age distributions of hospitalizations
and serious healthcare interventions administered to COVID-19
patients. In our study, we analysed known SARS-CoV-2 infection
records for the province of Ontario, and quantified the age-intervention associations
for hospitalization, ICU admission, intubations, and ventilations. We
quantified existing age-related impacts of COVID-19 and identified
potential future risks should the healthcare system become overwhelmed
with COVID-19 patients.

Ontario has so far been able to offer and provide COVID-19 treatment
to all that request it. Yet we find that the probability of survival
given hospitalization for COVID-19 decreases substantially after age
40. The distribution of COVID-19-related hospitalizations spans a much
wider age range than deaths, which are concentrated in the extremely
elderly. If prevalence of the virus were to increase in the general
population (as is currently occurring), and ICUs or hospitals were to
subsequently reach maximum capacity, the distribution of deaths might expand toward younger ages.

While our work is specific to the Ontario SARS-CoV-2 epidemic, it has
more general and important implications for COVID-19 outbreaks all
over the world, in terms of both public health and clinical
practice. Our study reveals a potentially serious mortality risk to
middle-aged individuals from SARS-CoV-2 infection if the healthcare
system were to be overwhelmed with COVID-19 patients; this risk is not
captured by the widely available distributions of known SARS-CoV-2
infections and COVID-19-related deaths.

We have presented our study to the Ontario government's COVID-19
science advisory panel (the Ontario Science Table,
\url{https://covid19-sciencetable.ca/}), where it was received with
great interest.  We believe that the readership of \journalname{}
should be made aware of our findings as soon as possible.

All authors have approved the manuscript for submission and declare no competing interests. A preprint of this work has been posted to medRxiv (\url{https://www.medrxiv.org/content/10.1101/2020.09.01.20186395v2}). The manuscript has not been published or submitted for publication elsewhere.

@@

Irena Papst, MSc
\setlength{\parskip}{0em}

PhD Candidate, Center for Applied Mathematics, Cornell University

@@

Michael Li, PhD

Postdoctoral Fellow, Department of Biology, McMaster University

Research Associate, South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch

@@

David Champredon, PhD

Postdoctoral Associate, Department of Pathology and Laboratory Medicine,

Western University

@@

@@

Professor, Department of Mathematics, University of Toronto
\end{flushleft}
\end{multicols}

>>>>

%% comment macros
\newcommand{\ali}[1]{\textcolor{blush}{$\langle$\textbf{AG:} #1$\rangle$}}

%% journal name macro
\newcommand{\journalname}{\emph{BMB}}
% Bulletin of Mathematical Biology
%% data macros
% \input{data_macros.tex}

@@

Dr. Alan Hastings and Dr. Reinhard Laubenbacher

@@

\begin{flushright}
Ali Gharouni

Postdoctoral Fellow

Department of Mathematics and Statistics,

McMaster University,

Hamilton, ON, Canada, L8S 4K1

\href{mailto:agharoun@uottawa.ca}{agharoun@uottawa.ca}
\end{flushright}

@@

Dear Drs.\ Hastings and Laubenbacher,

We are pleased to submit an original research article entitled
``Testing and Isolation Efficacy: Insights from a Simple Epidemic Model'' for your consideration for publication in \journalname{}. This study specifically aligns with the aims of \journalname{} to publish research related to the social determinants of health and adds new insight into the evolving COVID-19 pandemic.

We have developed a compartmental model to understand the effect of testing processes on epidemic dynamics. We studied the effects of testing intensity, rate of test return, and isolation efficacy, on epidemic dynamics when different levels of testing focus are in place. Our analysis shows that under some circumstances both increased testing intensity and faster test reporting can reduce the effectiveness of control, and allows us to explore the conditions under which this occurs. Conversely, we find that focusing testing on infected individuals always acts to increase effectiveness of control. Our work provides insights for epidemiologists, mathematical modelers and public health managers in the context of the interaction between testing processes and epidemic processes.

All authors have approved the manuscript for submission and declare no competing interests. A preprint of this work has been posted to arXiv (\url{https://arxiv.org/abs/2107.08259}). The manuscript has not been published or submitted for publication elsewhere.

@@

Ali Gharouni, PhD
\setlength{\parskip}{0em}

Postdoctoral Fellow, Department of Mathematics and Statistics, McMaster University

@@

Fady M. Abdelmalek, B.Sc.

Department of Mathematics and Statistics, McMaster University

@@

@@

\vfill\null

@@

South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch, South Africa

\end{flushleft}
\end{multicols}
